Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-22 paper/main.tex
add link to supplement
add short introduction for methodological work

% --------------------------------------------------------------------------------------------------
\subsection{Numerical Implementation}
We implemented the described model in our recently developed \ac{abm} framework
\texttt{cellular\_raza}~\cite{Pleyer2025}.
By design \texttt{cellular\_raza} does not assume any particular cellular representation, which
makes it flexible and easy to adapt the simulation to particular needs.
Its design allows us to experiment with different model implementations and facilitates the
explorative capabilities of researchers.
It is furthermore able to seamlessly integrate into Python code-bases thus allowing us to utilize
methods from other popular software libraries such as \texttt{numpy} and \texttt{scipy}.
Furthermore, it is an \ac{abm} which is able to natively track mother-daughter relations of cells
over the course of a numerical simulation which is an integral component of our imaging-based
comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).

%###################################################################################################
\section{Methodology}
\label{sec:methodology}
These sections introduce methods which can be combined in order to estimate the parameters of our
computational model.
It has to be noted that they only depend on the spatial representation described in
section~\ref{subsec:spatial-representation}.
Consequently, these methods can be applied to a broader class of mechanical models that may exhibit
varied cellular behaviours.

The developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate intracellular variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters between all agents for processes which can be expected to behave
identical within margin of error.
Furthermore, mixed approaches where some parameters are assigned globally and others individually
can be realized as well.

>>>>

% --------------------------------------------------------------------------------------------------
\subsection{Numerical Implementation}
We implemented the described model in our recently developed \ac{abm} framework
\texttt{cellular\_raza}~\cite{Pleyer2025}.
By design \texttt{cellular\_raza} does not assume any particular cellular representation, which
makes it flexible and easy to adapt the simulation to particular needs.
Its design allows us to experiment with different model implementations and facilitates the
explorative capabilities of researchers.
It is furthermore able to seamlessly integrate into Python code-bases thus allowing us to utilize
methods from other popular software libraries such as \texttt{numpy} and \texttt{scipy}.
Furthermore, it is an \ac{abm} which is able to natively track mother-daughter relations of cells
over the course of a numerical simulation which is an integral component of our imaging-based
comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).
Further details can be found in the supplement (see section~\ref{section:supplement-performance})

%###################################################################################################
\section{Methodology}
\label{sec:methodology}
These sections introduce methods which can be combined in order to estimate the parameters of our
computational model.
It is split into 3 parts, each covering distinct purposes.
We begin by explaining the visualization process which will later be used within our imaging-based
comparison algorithm (see section~\ref{subsec:imaging-based-comparison}) and for testing purposes.
Afterwards, we introduce a position extraction algorithm which extracts cellular information in the
form of collections of vertices from segmented microscopic images.
Finally, we will provide two distinct comparison algorithms which can be used to compare numerically
calculated results with experimental data.
We will test and benchmark the constructed algorithms within the coming sections.

It should be noted that these methods only depend on the spatial representation described in
section~\ref{subsec:spatial-representation}.
Consequently, these methods can be applied to a broader class of mechanical models that can exhibit
varied cellular behaviours.
Furthermore, the developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate intracellular variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters between all agents for processes which can be expected to behave
identical within margin of error.
Furthermore, mixed approaches where some parameters are assigned globally and others individually
can be realized as well.
